\batcave: Bayesian Analysis Tools for Context-Aware Variant Evaluation

### 2019-07-29 NAR_DRAFT.tex

\section{RESULTS}
% \begin{figure*}[t]
%   \begin{center}
%   \includegraphics{figures/aml_plot.pdf}
%   \end{center}
%   \caption{Model performance on 300X Whole leukemia genome.
%   \textbf{(a)} Precision/recall curves and \textbf{(c)} Reciever operating characteristic curves.
%   Distribution of estimated true positive probabilities for true positive (top) and true negative (bottom) variants for \textbf{b)} the MuTect model and \textbf{d)} the Calibrated model.
%   A perfectly calibrated model would generate the diagonal line.}
%   \label{NAR-aml}
% \end{figure*}
\begin{figure*}
  \begin{center}
  \includegraphics{figures/fig2.pdf}
  \end{center}
  \caption{Model performance on 500X simulated whole exome.
  \textbf{(a)} Precision recall curves and \textbf{(c)} Reciever operating characteristic curves for 3 mutation signatures.
  Distribution of estimated true positive probabilities for true positive (top) and true negative (bottom) variants for \textbf{b)} the MuTect model and \textbf{d)} the \batcave model.
  A perfectly calibrated model would generate the diagonal line.\bkmcomment{change the axis on b and d to be the same as the steyerberg paper. much clearer}}
\label{NAR-wes_fig}
\end{figure*}

@@

\begin{table*}[b]
\tableparts{%
\caption{This is a table with $\mu$, auroc, auprc, ICI, and maybe fraction called?}
\label{table:01}%
}{%
\begin{tabular*}{\textwidth}{@{}lllllllll@{}}
\toprule
Depth & Mutation profile & $\mu/\beta$ & AUROC & & AUPRC & & ICI & 
\\
& & (estimated) & MuTect & \batcave & MuTect & \batcave & MuTect & \batcave
\\
\colrule
100X whole genome & Concentrated & 3.6e-7 & Row 1 & -- & -- & -- & -- & --
\\
100X whole genome & Intermediate & 3.2e-7 & Row 1 & -- & -- & -- & -- & --
\\
100X whole genome & Diffuse & 3.2e-7 & Row 1 & -- & -- & -- & -- & --
\\
500X whole exome & Concentrated & 3.6e-7 & .848 & .928 & .674 & .757 & .0585 & .0465
\\
500X whole exome & Intermediate & 3.6e-7 & .847 & .842 & .677 & .707 & .0587 & .0566 
\\
500X whole exome & Diffuse & 3.6e-7 & .850 & .873 & .676 & .698 & .0565 & .0575
\\
AML 31 & Actual & 3.6e-8 & .935 & .949 & .995 & .996 & 0.121 & 0.155
\\
\botrule
\end{tabular*}%
}
{$\mu$ = per-base mutation rate, AUROC/AUPRC = area under ROC/PRC curve, ICI = Integrated calibration index.}
\end{table*}

@@

Simulations suggest that if the mutation profile is the data generating distribution, \batcave can estimate effective mutation rate and data generating distribution, and doing so improves both classification and calibration.

\subsection{Discussion subsection one}

@@

>>>>

\section{RESULTS}
% \begin{figure*}[t]
%   \begin{center}
%   \includegraphics{figures/aml_plot.pdf}
%   \end{center}
%   \caption{Model performance on 300X Whole leukemia genome.
%   \textbf{(a)} Precision/recall curves and \textbf{(c)} Reciever operating characteristic curves.
%   Distribution of estimated true positive probabilities for true positive (top) and true negative (bottom) variants for \textbf{b)} the MuTect model and \textbf{d)} the Calibrated model.
%   A perfectly calibrated model would generate the diagonal line.}
%   \label{NAR-aml}
% \end{figure*}
\begin{figure*}
  \begin{center}
  \includegraphics{figures/fig_wes.pdf}
  \end{center}
  \caption{Model performance on 500X simulated whole exome.
  \textbf{(a)} Precision recall curves and \textbf{(c)} Reciever operating characteristic curves for 3 mutation signatures.
  Distribution of estimated true positive probabilities for true positive (top) and true negative (bottom) variants for \textbf{b)} the MuTect model and \textbf{d)} the \batcave model.
  A perfectly calibrated model would generate the diagonal line.\bkmcomment{change the axis on b and d to be the same as the steyerberg paper. much clearer}}
\label{NAR-wes_fig}
\end{figure*}

@@

\begin{table*}[b]
\tableparts{%
\caption{This is a table with $\mu$, auroc, auprc, ICI, and maybe fraction called?}
\label{table:01}%
}{%
\begin{tabular*}{\textwidth}{@{}lllllllll@{}}
\toprule
Depth & Mutation profile & $\mu/\beta$ & AUROC & & AUPRC & & ICI & 
\\
& & (estimated) & MuTect & \batcave & MuTect & \batcave & MuTect & \batcave
\\
\colrule
100X whole genome & Concentrated & 3.6e-7 & .987 & .993 & .972 & .975 & -- & --
\\
100X whole genome & Intermediate & 3.2e-7 & .987 & .989 & .972 & .973 & -- & --
\\
100X whole genome & Diffuse & 3.2e-7 & .988 & .989 & .971 & .973 & -- & --
\\
500X whole exome & Concentrated & 3.6e-7 & .848 & .929 & .674 & .758 & .0585 & .0462
\\
500X whole exome & Intermediate & 3.6e-7 & .847 & .881 & .677 & .706 & .0587 & .0568 
\\
500X whole exome & Diffuse & 3.6e-7 & .850 & .873 & .676 & .698 & .0565 & .0576
\\
AML 31 & Actual & 3.6e-8 & .935 & .949 & .995 & .996 & 0.121 & 0.155
\\
\botrule
\end{tabular*}%
}
{$\mu$ = per-base mutation rate, AUROC/AUPRC = area under ROC/PRC curve, ICI = Integrated calibration index.}
\end{table*}

@@

-- Simulations suggest that if the mutation profile is the data generating distribution, \batcave can estimate effective mutation rate and data generating distribution, and doing so improves both classification and calibration.

-- One hope is to enable higher confidence in low frequency alleles as sequencing depth increases. Depth seems to have a very large effect on performance i.e. results are more compelling for 500X wes than for 100x wgs, although if you do 100 wgs you will still benefit.
In other words sequence depth matters more performance wise than sequencing breadth if the only desire is to see low frequency alleles. No advantage is gained (after a certain depth) of adding variants(sites) by wgs sequencing.
\subsection{Discussion subsection one}

@@

\section{Supplementary Figures}

\begin{figure*}
  \begin{center}
  \includegraphics{figures/fig_wes.pdf}
  \end{center}
  \caption{Model performance on 100X simulated whole genome.
  \textbf{(a)} Precision recall curves and \textbf{(c)} Reciever operating characteristic curves for 3 mutation signatures.
  Distribution of estimated true positive probabilities for true positive (top) and true negative (bottom) variants for \textbf{b)} the MuTect model and \textbf{d)} the \batcave model.
  A perfectly calibrated model would generate the diagonal line.\bkmcomment{change the axis on b and d to be the same as the steyerberg paper. much clearer}}
\label{NAR-wgs_fig}
\end{figure*}
